---
title: "Workshop on Data Visualization in R"
author: "Karthy"
date: "2/25/2020"
output:
  slidy_presentation:
    keep_md: yes
  ioslides_presentation: default
  powerpoint_presentation: default
  beamer_presentation: 
    fonttheme: professionalfonts
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Visualization in R

- One Variable Visualization 


- Two Variable Visualization 



- Correlation Relationship 


- Regression Analysis Relationship


## Loading major libraries
These are the basic libraries in R visualization.
```{r  echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)

library(lattice)

library(latticeExtra)

library(ggpubr)

library(scatterplot3d)

library(gridExtra)

library(ggfortify)

##shiny::runGitHub("cardiomoon/ggplot2new")
```

## Loading major libraries
```{r  echo=TRUE, message=FALSE, warning=FALSE}
library(ggmosaic)

library(ggeffects)

library(GGally)

library(ggcorrplot)

library(poliscidata)

library(wooldridge)

library(easystats)
```


## One Variable Visualization

- For continuous variable: histogram, frequency plot, and density plots
- Examples: Age, income, gdp....
  
  
  
- For Discrete Variable:bargraph
  
- Examples: gender, race, conflict, civil war.....
  
## R Code for One Variable Visualization
```{r R code, echo=TRUE, message=FALSE, warning=FALSE}

## Histogram in ggplot package
ggplot(data = diamonds, aes(price)) + geom_histogram()

## Histogram varied based on the cut of the diamonds (group)
ggplot(data = diamonds, aes(price, color = cut)) + geom_histogram()


## Densityplot in ggplot package
ggplot(data = diamonds, aes(price)) + geom_density()

## Densityplot varied based on the cut of the diamonds (group)
ggplot(data = diamonds, aes(price, color = cut)) + geom_density()

## Frequency plot in ggplot package
ggplot(data = diamonds, aes(price)) + geom_freqpoly()

## Frequency plot is visualized based on the cut (group)
ggplot(data = diamonds, aes(price, color = cut)) + geom_freqpoly()


## Histogram in lattice package. Price is visualized based on the cut of the diamonds. 
histogram(~ price | factor(cut), data = diamonds)

## Histogram using ggpubr package. Price is visualized based on cut of the diamonds.
gghistogram(diamonds, x = "price",
            add = "mean", rug = TRUE, color = "cut",
            palette = c("red", "blue", "green", "brown", "violet"))

## Densityplots using ggpubr package. Price is visualized based on cut of the diamonds.
ggdensity(diamonds, x = "price",
          add = "mean", rug = TRUE,
          color = "cut",
          palette = c("red", "blue", "green", "brown", "violet"))



```




## R Code for Discrete One Variable Visualization
```{r Onevar, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data = world, aes(dem_level4, fill = gdp_cap3)) + geom_bar() 

       
ggplot(data = world, aes(dem_level4, fill = regime_type3)) + geom_bar()

## Facet Grid attachment to separate based on different regions
ggplot(data = world, aes(dem_level4, fill = regime_type3)) + geom_bar() + facet_grid(regionun ~.)



```

## R Code for Two-variable visualization 
```{r Twovar, echo=TRUE, message=FALSE, warning=FALSE}
## Discrete variable (X) and continous variable (Y)
## Boxplot Visualization. Gini Coefficient (inequality) (y-variable) visualized based on regime types (x-variable). 
ggplot(data = world, aes(x = regime_type3, y = gini10, color = oecd)) + geom_boxplot() 

ggplot(data = world, aes(x = dem_level4, y =  gini10, color = oecd)) + geom_boxplot() + coord_flip()


bwplot(factor(dem_level4) ~ gini10 , data = world, xlab = "Gini Coefficient", col = "green", fill  = "blue")


## Using the ggpubr package. 
ggbarplot(data = world, x =  "dem_level4", y = "gini10", fill = "regime_type3", xlab = "Democracy Level", ylab = "Gini Coefficient", short.panel.labs= T)



## Continuous variables on both axes. Levels of gini-coefficient (y) is visualized based on democracy score (x)

ggplot(data = world, aes(x = dem_score14, y = gini10)) + geom_point()  

ggplot(data = world, aes(x = dem_score14, y = gini10, color = gdp_cap3)) + geom_point() 

ggplot(data = world , aes(x = dem_score14, y = gini10, color = gdp_cap3)) + geom_point() + facet_wrap(~regionun)

## using Lattice package 
xyplot(gini10~dem_score14 | regime_type3 , data = world, 
       type = "h", xlab = "Democracy Levels")


## Discrete variable (X) and Discrete variable (Y). Here regime type is visualized with GDP per capita. 
ggplot(data = world, aes(x = dem_level4 , y = gdp_cap3, color = regime_type3)) + geom_count()  







```


## R Code for Correlation Relationships

```{r Corrcode, echo=TRUE, message=FALSE, warning=FALSE}
##. Correlation of numerical variables
corr = cor(diamonds[, c(1, 5:7)])
ggcorrplot(corr = corr)

## using GGally package. The Bush approval ratings is used to visualize the possible correlations. 
ggpairs(approval, columns = 7:11, ggplot2::aes(color=factor(X11.Sep)))

```


##. R Code for Regression Relationships

```{r reg, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
##using ggplot package for regression line fit
ggplot(data = world, aes(x = dem_score14, y = gini10 )) + geom_point(na.rm = T) + geom_smooth(method = "lm", na.rm = T)  

##using ggplot package for regression line fit (loess method)
ggplot(data = world, aes(x = dem_score14, y = gini10)) + geom_point(na.rm = T) + geom_smooth(method = "loess", na.rm = T) 

##using quantile regression estimation
ggplot(data = world, aes(x = dem_score14, y = gini10)) +geom_point(na.rm = T) + geom_quantile(na.rm = T)


```

## Regression diagnostics using ggpubr and ggfortify 
```{r regdiag, echo=TRUE, message=FALSE, warning=FALSE}
## Using ggpubr package
ggscatter(data = world, x = "dem_score14", y = "gini10",
   color = "black", shape = 21, size = 3, # Points color, shape and size
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")
   )

##using ggeffects, ggfortify, and performance package to estimate and diagnosize the linearmodels
##Model.1: predicting and diagonizing the generalized linear model
m1 = lm(approve~gasprice + unemploy + katrina + lrgasprice, data = approval)
autoplot(m1, which = 1:6, ncol = 3, label.size = 3)


##Model.2: predicting and diagonizing the generalized linear model with addition of iraq invasion variable
m2 = lm(approve~gasprice + unemploy + katrina + lrgasprice + iraqinvade, data = approval)
autoplot(m2, which = 1:6, ncol = 3, label.size = 3)

##Model.3: predicting and diagonizing the generalized linear model with control variables
m3 = lm(approve~gasprice + unemploy + katrina + lrgasprice + iraqinvade + lcpifood + X11.Sep, data = approval)
autoplot(m3, which = 1:6, ncol = 3, label.size = 3)

## Coefficient estimates for various models
g1 = ggcoef(m1, vline_color = "green", sort = "ascending", exclude_intercept = T)
g2 = ggcoef(m2, vline_color = "blue", sort = "ascending", exclude_intercept = T)
g3 = ggcoef(m3, vline_color = "red", sort = "ascending", exclude_intercept = T)

grid.arrange(g1, g2, g3, nrow=3)
```


